---
title: "EU-Skepsis und Wahlbeteiligung bei Wahlen zum europäischen Parlament"
author: "A. Seitz"
output: html_notebook
---

*This notebook contains all steps necessary to arrive at the three regression models presented in the thesis.*
*Running all chunks in this notebook results in a file "results.RData" which is required to compile the thesis document.*

# Clear workspace

```{r echo=FALSE}
rm(list = ls())
```

# Load libraries

```{r echo=FALSE}
library(tidyverse)
library(foreign)
library(lubridate)
library(lme4)
library(sjlabelled)
library(sjPlot)
library(jtools)
library(rlang)
```

# Load datasets

## Official turnout
```{r echo=FALSE}
countries <- read.csv("data/turnout.csv")
```

## European Election Study Post-Election Cross-Section 2019
```{r echo=FALSE, warning=FALSE}
ees2019.raw.lab <- read.spss("data/ees2019.sav", to.data.frame = TRUE)
ees2019.raw <- read.spss("data/ees2019.sav", to.data.frame = TRUE, use.value.labels = FALSE)
ees2019.raw$hCountry <- as.character(ees2019.raw.lab$hCountry)
ees2019.raw <- 
  ees2019.raw %>% 
  mutate(hCountry = recode(hCountry, "UK" = "United Kingdom")) %>% 
  remove_all_labels()
rm(ees2019.raw.lab)
```

## ParlGov
```{r echo=FALSE, warning=FALSE}
pg <- read.csv("data/parlgov_view_election.csv") %>%
  mutate(election_date = ymd(election_date))

pg.dates.nat <-
  pg %>%
  filter(election_type == "parliament") %>%
  select(Country = country_name, Date = election_date) %>%
  mutate(Year = year(Date)) %>%
  unique()

pg.dates.ep <-
  pg %>%
  filter(election_type == "ep") %>%
  select(Country = country_name, Date = election_date) %>%
  mutate(Year = year(Date)) %>%
  unique()

pg.ep <- 
  pg %>%
  filter(election_type == "ep") %>%
  mutate(Year = year(election_date)) %>%
  filter(Year >= 2004) %>%
  inner_join(
    read.csv("data/parlgov_view_party.csv") %>% select(party_id, eu_anti_pro))
```

## PopuList
```{r echo=FALSE, warning=FALSE}
populist <- read.csv("data/PopuList.csv")
```

## Additional member state level variables
```{r}
addvars_countries <- read.csv("data/addvars_countries.csv")
```

# Data Processing

## Calculate eurosceptic vote shares
```{r}
euroscep <- 
  left_join(
    populist %>% rename(party_id = parlgov_id),
    pg.ep,
    by = "party_id") %>%
  filter((eurosceptic_startnobl <= Year & eurosceptic_endnobl > Year) | eu_anti_pro < 5) %>%
  mutate(vote_share = vote_share %>% replace_na(0)) %>%
  rename(Country = country_name.x) %>%
  group_by(Country, Year) %>%
  summarise(EuroscepticVoteShare = sum(vote_share) / 100)

countries <-
  left_join(countries, euroscep)

countries$EuroscepticVoteShare[is.na(countries$EuroscepticVoteShare)] <- 0
```

## Calculate "electoral salience" (time since previous national parliament election)
```{r}
countries <- inner_join(countries, pg.dates.ep)

last_national_election_date <- function(country, date) {
  res <- pg.dates.nat %>% filter(Country == country & Date < date)
  max(res$Date)
}

countries <- 
  countries %>% 
  mutate(last_nat = as_date(mapply(last_national_election_date, Country, Date))) %>%
  mutate(YearsSinceLastNationalElection = as.numeric((Date - last_nat) / 365.25))
```

## Calculate number of Previous EP elections
```{r}
nPrev <- function(country, year) {
  nrow(countries %>% filter(Country == country & Year < year & IsNat == 0))
}

countries <- countries %>%
  mutate(PreviousElections = mapply(nPrev, Country, Year))
```

## Check for sunday voting
```{r}
countries <- countries %>% mutate(Sunday = as.factor(wday(Date) == 1))
```

## Remove duplicate rows for Bulgaria and Romania 2007
```{r}
countries <-
  countries %>% filter((Country != "Bulgaria" & Country != "Romania") | Date != ymd("2007-01-01"))
```

## EES 2019 data recoding
This is mostly to invert scales and change factor levels, and also to give variables intelligible names.
```{r echo=FALSE}
ees2019 <- 
  ees2019.raw %>%
  filter(Q6 <= 2) %>%
  select(
    Country = hCountry, 
    DSN = Q3, 
    DSEU = Q4, 
    GovS = Q5,
    Voted = Q6,
    CA = Q8,
    LR = Q11, 
    PTN = Q18_1, 
    PTEU = Q18_2, 
    Econ = Q19, 
    EUM = Q22, 
    Unif = Q23,
    PI = Q21, 
    Party = Q26, 
    Union = D1, 
    Edu = EDU,
    Gender = D3, 
    Age = D4_1,
    Partner = D5,
    Empl = D6, 
    Class = D7, 
    Urban = D8, 
    Relig = D10
    ) %>%
  na_if(98) %>%
  na_if(97) %>%
  na_if(99) %>%
  mutate(Country = as.factor(Country)) %>%
  mutate(DSN = 4 - DSN) %>%
  mutate(DSEU = 4 - DSEU) %>%
  mutate(GovS = GovS %>% replace(GovS == 2, 0)) %>%
  mutate(Voted = Voted %>% replace(Voted == 2, 0)) %>%
  mutate(PTN = 5 - PTN) %>%
  mutate(PTEU = 5 - PTEU) %>%
  mutate(Econ = 5 - Econ) %>%
  mutate(PI = 4 - PI) %>%
  mutate(EUM = EUM %>% 
    replace(EUM == 2, 0) %>% 
    replace(EUM == 1, 2) %>%
    replace(EUM == 3, 1) %>% 
    as.factor()) %>%
  mutate(Party = (4 - Party) %>% replace(Party == 4, 0)) %>%
  mutate(Union = Union %>%
    replace(Union == 4, 0) %>%
    replace(Union == 2, 1) %>% 
    replace(Union == 3, 2) %>% 
    replace(Union == 1, 2) %>% 
    as.factor()) %>%
  mutate(Gender = Gender %>% as.factor()) %>%
  mutate(Age = 2019 - Age) %>%
  mutate(Partner = as.numeric(Partner <= 8)) %>%
  mutate(Empl = Empl %>% 
    replace(Empl < 4, 1) %>%
    replace(Empl >= 4, 0) %>%
    replace(Empl == 1, 0)) %>%
  mutate(Class = Class %>% replace(Class == 6, NA)) %>%
  mutate(Relig = as.numeric(Relig <= 3))
```

## Clean up country-level dataset
```{r}
countries2019 <-
  countries %>% 
  filter(Year == 2019) %>% 
  select(Country, C_EUSkep = EuroscepticVoteShare, C_ElecSal = YearsSinceLastNationalElection, C_Prev = PreviousElections, C_Sun = Sunday, Turnout) %>%
  left_join(addvars_countries %>% select(Country, C_Comp = CompulsoryVoting, C_Simul = Simul)) %>%
  mutate(C_Sun = as.integer(C_Sun))
```

## Merge the EES 2019 (individual-level) data with the Country (macro-level) data
```{r}
data.full <- left_join(ees2019, countries2019 %>% select(-Turnout))
```

## Drop rows with missing values
```{r}
data <- data.full %>% select(Country, Voted, DSEU, EUM, Unif, PTEU, DSN, PTN, Econ, GovS, CA, PI, Party, Union, Edu, Gender, Age, Partner, Empl, Class, Urban, Relig, C_EUSkep, C_ElecSal, C_Prev, C_Comp, C_Simul, C_Sun) %>% drop_na()
```

## Set factor response labels
```{r}
data$EUM = set_labels(data$EUM, labels = c("schlecht", "neutral", "gut"))
data$Union = set_labels(data$Union, labels = c("keine", "direkt", "indirekt"))
data$Gender = set_labels(data$Gender, labels = c("m", "w", "d"))
```

# Models

## Model 0: Multilevel model with controls only
```{r}
mod0 <- glmer(
  Voted ~ 
  DSN + PTN + Econ + GovS +
  CA + PI + Party + Union + Edu + Gender + Age + Partner + Empl + Class + Urban + Relig +
  C_ElecSal + C_Prev + C_Comp + C_Simul + C_Sun +
  (1 | Country),
  data = data, family = "binomial") 
summ(mod0, confint = TRUE)
```

## Model 1: Multilevel model with all predictors, random intercept only
```{r warning=FALSE}
mod1 <- glmer(
  Voted ~ 
  EUM + Unif + PTEU + DSEU + PTN + DSN + Econ + GovS +
  CA + PI + Party + Union + Edu + Gender + Age + Partner + Empl + Class + Urban + Relig +
  C_EUSkep + C_ElecSal + C_Prev + C_Comp + C_Simul + C_Sun +
  EUM:C_EUSkep + Unif:C_EUSkep + PTEU:C_EUSkep + DSEU:C_EUSkep + 
  (1 | Country),
  data = data, family = "binomial") 
summ(mod1, confint = TRUE)
```

## Model 2: Multilevel model with all predictors and random slopes for EU attitude variables
```{r warning=FALSE}
mod2 <- glmer(
  Voted ~ 
  DSEU + EUM + Unif + PTEU + DSN + PTN + Econ + GovS +
  CA + PI + Party + Union + Edu + Gender + Age + Partner + Empl + Class + Urban + Relig +
  C_EUSkep + C_ElecSal + C_Prev + C_Comp + C_Simul + C_Sun +
  DSEU:C_EUSkep + EUM:C_EUSkep + Unif:C_EUSkep + PTEU:C_EUSkep +
  (1 + DSEU + EUM + Unif + PTEU | Country),
  data = data, family = "binomial") 
summ(mod1, confint = TRUE)
```

## Model comparison: AIC/BIC, likelihood ratio test
```{r}
mod_comp <- anova(mod0, mod1, mod2)
mod_comp
```

## Country-level OLS model
I ended up not using this in the paper at all, but decided to keep it in here for transparency.
```{r}
mod_c <- lm(Turnout ~ C_EUSkep + C_ElecSal + C_Prev + C_Comp + C_Simul + C_Sun, data = countries2019) %>% scale_mod()
summ(mod_c)
```

# Save calculation results for use in the thesis document

```{r echo=FALSE}
save.image(file = "results.RData")
```
